Periodograms for Multiband Astronomical Time Series

### 2015-01-30 writeup/paper.tex
Add LCB14 discussion

\section{A Multiple-Band Model}
\sectlabel{multiband}
In this section we will apply what was learned in the previous sections to construct the {\it multiband periodogram} which can flexibly account for heterogeneous sources of data for a single object.
To compute such a periodogram, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
The model contains the following features:
\begin{enumerate}
  \item An $N_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual of a single band from this overall variability.
\end{enumerate}
The total number of parameters for $K$ filters is then $M_K = (2N_{base} + 1) + K(2N_{band} + 1)$. As a result, for each band $k$ we have the following model of the observed magnitudes:
\begin{eqnarray}
  y_k(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\\ 
  &\theta^{(k)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(k)}_{2n - 1}\sin(n\omega t) + \theta^{(k)}_{2n}\cos(n\omega t)\right].&
\end{eqnarray}
The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(k)}$ are determined individually.

@@

>>>>

\section{A Multiple-Band Model}
\sectlabel{multiband}
In this section we will apply what was learned in the previous sections to construct the {\it multiband periodogram} which can flexibly account for heterogeneous sources of data for a single object.
To compute such a periodogram, we will take advantage of the easy extensibility of the matrix formalism which led to our generalizations above.
The model contains the following features:
\begin{enumerate}
  \item An $N_{base}$-term truncated Fourier fit which models a latent parameter, which we'll call the ``overall variability''.
  \item A set of $N_{band}$-term truncated Fourier fits, each of which models the residual of a single band from this overall variability.
\end{enumerate}
The total number of parameters for $K$ filters is then $M_K = (2N_{base} + 1) + K(2N_{band} + 1)$. As a result, for each band $k$ we have the following model of the observed magnitudes:
\begin{eqnarray}
  \eqlabel{multiband_model}
  y_k(t|\omega,\theta) = &\theta_0 + \sum_{n=1}^{M_{base}} \left[\theta_{2n - 1}\sin(n\omega t) + \theta_{2n}\cos(n\omega t)\right]& +\nonumber\\ 
  &\theta^{(k)}_0 + \sum_{n=1}^{M_{band}} \left[\theta^{(k)}_{2n - 1}\sin(n\omega t) + \theta^{(k)}_{2n}\cos(n\omega t)\right].&
\end{eqnarray}
The important feature of this model is that {\it all bands} share the same base parameters $\theta$, while their offsets $\theta^{(k)}$ are determined individually.

@@

\subsection{Comparison with \citet{Long2014}}
As we were finishing this study, it was pointed out to us that another group had released a preprint independently addressing the multiband periodogram case \citep[][hereafter LCB14]{Long2014}, and come up with a solutions very similar to the ones presented here.
They present two methods, the ``Multiband Generalized Lomb-Scargle'' (MGLS) which is effectively identical to the $(1, 0)$ multi-phase model, and the ``Penalized Generalized Lomb-Scargle'' (PGLS), which is similar in spirit to our $(0, 1)$ shared-phase model.

In the PGLS model, they start with a multi-phase model, fitting independent $N=1$ term fits to each band, and apply a nonlinear regularization term which penalizes differences in the amplitude and phase. In terms of the formalism used in this work, the PGLS model minimizes a regularized $\chi^2$ of the form
\begin{equation}
  \chi^2_{PGLS} = \sum_{k=1}^K \bigg[~\chi^2_{GLS}(D^{(K)}) + J_A(A^{(k)}) + J_\phi(\phi^{(k)})~\bigg].
\end{equation}
for $K$ bands, where $\chi^2_{GLS}(D^{(K)})$ is the $\chi^2$ of the standard floating mean model on the single-band data $D^{(K)}$, and $J_A$ and $J_\phi$ are regularization/penalty terms which are a function of the amplitude $A^{k}$ and phase $\phi^{(k)}$ of each model. In terms of our linear model parameters $\theta^{(k)}$, this amplitude and phase can be expressed:
\begin{eqnarray}
 A^{(k)} &=& \sqrt{(\theta_1^{(k)})^2 + (\theta_1^{(k)})^2}\nonumber\\
 \phi^{(k)} &=& \arctan(\theta_2 / \theta_1)
\end{eqnarray}
The selected form of these regularization terms penalizes deviations of the amplitude and phase from a common mean between the bands; in this sense the PGLS model can be considered a conceptual mid-point between our shared-phase and multi-phase models. Within the formalism proposed here, such a mid-point may be alternatively attained by suitably increasing the regularization parameter $\epsilon$ used in the shared-phase model, though the nature of the resulting regularization will differ.

Computationally, the PGLS model requires a nonlinear optimization at each frequency $\omega$, and is thus much more expensive than the straightforward linear optimization of our shared-phase model.
For this reason, LCB14 propose a clever method by which nested models are used to reduce the number of nonlinear optimizations used: essentially, by showing that the (linear) MGLS $\chi^2$ is a lower-bound of the (non-linear) PGLS $\chi^2$, it is possible to iteratively reduce the number of PGLS computations required to minimize the $\chi^2$ among a grid of frequencies.
Such an optimization could also be applied in the case of our shared-phase model, but is not necessary here due to its relative speed.
When applying the method to a very large number of light curves in, e.g. LSST however, such a computational trick may prove very useful.

Given these important distinctions between the models proposed here and in LCB14, in future work we plan to do a detailed comparison of the two means of model regularization.
